---
title: "Final_Project_Draft"
author: "Eldonna Weber"
date: "5/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Intro

## What:  
For my final project, I analyzed Zillow home sales data for Mound,
MN over the past two years, specially focusing on 3 bedroom, 2+ bathroom houses.

## Why:  
I have a rental house in Mound, MN that I'm planning to put on the market in 
June and I'm interested to see if the past home sales data will reflect any 
interesting trends or help me predict the final sale price of my house.

## How:  
I started by manually curating a data frame ('moundsales') from Zillow 
[(Data Source)](https://www.zillow.com/mound-mn/sold/house_type/3-3_beds/2.0-_baths/3_p/?searchQueryState=%7B%22pagination%22%3A%7B%22currentPage%22%3A3%7D%2C%22usersSearchTerm%22%3A%22Mound%2C%20MN%22%2C%22mapBounds%22%3A%7B%22west%22%3A-93.75159852978517%2C%22east%22%3A-93.56963747021486%2C%22south%22%3A44.87695033358592%2C%22north%22%3A44.992152656806155%7D%2C%22mapZoom%22%3A12%2C%22regionSelection%22%3A%5B%7B%22regionId%22%3A26000%2C%22regionType%22%3A6%7D%5D%2C%22isMapVisible%22%3Atrue%2C%22filterState%22%3A%7B%22beds%22%3A%7B%22max%22%3A3%2C%22min%22%3A3%7D%2C%22baths%22%3A%7B%22min%22%3A2%7D%2C%22con%22%3A%7B%22value%22%3Afalse%7D%2C%22fore%22%3A%7B%22value%22%3Afalse%7D%2C%22apa%22%3A%7B%22value%22%3Afalse%7D%2C%22mf%22%3A%7B%22value%22%3Afalse%7D%2C%22ah%22%3A%7B%22value%22%3Atrue%7D%2C%22sort%22%3A%7B%22value%22%3A%22globalrelevanceex%22%7D%2C%22nc%22%3A%7B%22value%22%3Afalse%7D%2C%22rs%22%3A%7B%22value%22%3Atrue%7D%2C%22land%22%3A%7B%22value%22%3Afalse%7D%2C%22tow%22%3A%7B%22value%22%3Afalse%7D%2C%22manu%22%3A%7B%22value%22%3Afalse%7D%2C%22cmsn%22%3A%7B%22value%22%3Afalse%7D%2C%22apco%22%3A%7B%22value%22%3Afalse%7D%2C%22fsba%22%3A%7B%22value%22%3Afalse%7D%7D%2C%22isListVisible%22%3Atrue%7D).

The data frame includes 150 instances and 12 variables:  'ID', 'Address', 
'Beds', 'Baths', 'SQFT', 'Month.Sold', 'Year.Sold', 'Sale.Price', 'Lake.Front', 
'Year.Built', 'Lot.Size.Sqft', 'Most.Recent.Tax.Assessment'.

I conducted exploratory data analysis of the data frame, created new 
data subsets, summarized data via histogram, scatter and box plots, determined 
probability, determined mean/standard deviation/z-score, and conducted 
regression.

  

# Body

According to Zillow, the were 150 three bedroom, two-plus bathroom houses sold 
in Mound, MN between April 2020 and April 2022. Mound is a community on 
Lake Minnetonka, therefore 32 of the 150 houses are lake front properties.  This
is an important distinction because these houses sold for significantly more 
than non-lake front houses, usually in the million dollar range. 

My house was built in 1948, has 1859 finished sqft and includes three bedrooms
and two baths.  The house sits on a 10,000 sqft, non-lake front, lot and was 
assessed at $245,000 in 2021. 

I'd like to understand the relationship between sale price and the most recent
tax assessment data and determine the probability of my house selling for over 
$300,000.



# Topics From Class

## Topic 1: Data Basics

For this observational study, I created a data frame labeled 'moundsales'.  In 
addition, I created a subset labeled 'nonlf' that only includes non-lake front
houses because including lake front houses skews the sale price and tax 
assessment data.

```{r}
moundsales <- read.table("moundsales.csv", sep = ",", header = TRUE)
```


```{r}
nonlf <- subset(moundsales, moundsales$Lake.Front == 'N')
```

I conducted a number of various data exploration activities in R (e.g., dim, 
length, names, summary, head, tail, etc.) to anlayze both the 'moundsales' and
'nonlf' data frames.  See Appendix for more details.


## Topic 2:  Summarizing Data

The mean sale price for lake front houses was $888,192, while the mean sale 
price for non-lake front houses is $324,801.

```{r}
lakefront <- subset(moundsales, moundsales$Lake.Front == 'Y')
summary(lakefront)
boxplot(moundsales$Sale.Price ~ moundsales$Lake.Front)
```
    

Only 46 houses, or 39%, sold for under $300,000, which supports my realator's 
claim that there has been a shortage of houses under $300,000 for sale in 
Mound, MN.


```{r}
under300k <- subset(nonlf, nonlf$Sale.Price < 300000)
dim(under300k)
```


Of the 46 houses that sold for more than $300,000, 16 had a recent tax 
assessment of $245,000 or below.


```{r}
S300kT245 <- subset(nonlf, nonlf$Sale.Price > 300000 & nonlf$Most.Recent.Tax.Assessment < 246000)
dim(S300kT245)
```




## Topic 3: Probability


The probability of a three bedroom, two-plus bathroom house in Mound, MN selling
for $300,000 or more given it is not lake front is 61%.

```{r}
Over299k <- subset(nonlf, nonlf$Sale.Price > 299999)
72/118
```


 
The probability of a three bedroom, two-plus bathroom house in Mound, MN selling
for $1 million or more given it has lake front is 28%.
```{r}
table(moundsales$Lake.Front)
```

```{r}
milsale <- subset(moundsales, moundsales$Sale.Price > 999999)
View(milsale)
```

```{r}
9/32
```



## Topic 4: Normal Distribution

The sale price distribution for the 'nonlf' data set is unimodal and normal with
a slight skew to the right.  The distribution has the following statistics:

mean = 324801
sd = 74870
x = 300000
Z-score =  - .331

There is a 63% probability that my house will sell for over $300k.



```{r}
format(nonlf$Sale.Price, scientific = FALSE)
hist(nonlf$Sale.Price)
```

```{r}
sd(nonlf$Sale.Price)
```


Given: 

mean = 324801
sd = 74870
x = 300000

```{r}
(300000 - 324801)/74870
```
Answer:  z-score = - .331


```{r}
1 - pnorm(-.331)
```
Answer:  63%



## Topic 5: Regression


According to the regression data below, there is a very high correlation between
the 'Sale.Price' and 'Most.Recent.Tax.Assessment' data.  The intercept value 
indicates that one could expect a house to sell for a premium of $20,934 over
the most recent tax assessment for the house.



```{r}
plot(moundsales$Sale.Price ~ moundsales$Most.Recent.Tax.Assessment)
lm(moundsales$Sale.Price ~ moundsales$Most.Recent.Tax.Assessment)
summary(lm(moundsales$Sale.Price ~ moundsales$Most.Recent.Tax.Assessment))
```


# Conclusion

In conclusion, I determined that there is a very high correlation between sale 
price and the most recent tax assessment for 3 bedroom, 2+ bath houses sold in 
Mound, MN in the past two years.  During this time, 72 non-lake front houses 
were sold for $300,000 or more and there is a 63% probability that my house will 
sell for $300,000 or more.  As a side note, I also learned that if my property
were to have lake front, there would be a 28% probability of selling my property
for $1 million or more!






# APPENDIX

### 'moundsales' Data Set

```{r}
moundsales <- read.table("moundsales.csv", sep = ",", header = TRUE)
```


```{r}
dim(moundsales)
length(dim(moundsales))
names(moundsales)
```

```{r}
head(moundsales)
tail(moundsales)
```

```{r}
summary(moundsales$Sale.Price)
summary(moundsales$Most.Recent.Tax.Assessment)
```
```{r}
table(moundsales$Lake.Front)
```

```{r}
milsale <- subset(moundsales, moundsales$Sale.Price > 999999)
View(milsale)
```


```{r}
lakefront <- table(moundsales$Lake.Front)
barplot(lakefront)
```
### 'nonlf' Data Set

```{r}
nonlf <- subset(moundsales, moundsales$Lake.Front == 'N')
```


```{r}
summary(nonlf$Sale.Price)
summary(nonlf$Most.Recent.Tax.Assessment)
```

```{r}
summary(nonlf$SQFT)
```

```{r}
summary(nonlf$Lot.Size.Sqft)
```
```{r}
summary(nonlf$Year.Built)
```

```{r}
under300k <- subset(nonlf, nonlf$Sale.Price < 300000)
dim(under300k)
```




